### 2016-12-01 report/figures/local-delta.tex

\begin{tikzpicture}
\begin{groupplot}[
group style={
	group size=3 by 1,
	vertical sep=0pt,
	horizontal sep=4mm,
	xlabels at=edge bottom,
	ylabels at=edge left,
	xticklabels at=edge bottom,
	yticklabels at=edge left,
},
every axis plot/.append style={very thick},
width=\linewidth / 2.5,
grid=major,
grid style={dashed,gray!30},
ymax=1.05,
ymin=0,
xmax=3.5,
xlabel={Time $\left[\si{\second}\right]$},
x tick label style={/pgf/number format/fixed},
ytick={0,0.2,0.4,0.6,0.8,1},
yticklabels={0,0.2,0.4,0.6,0.8,1},
ylabel=Percentiles,
legend columns=3,
legend cell align=left,
legend style={at={(1.7,1.4)},anchor=north, font=\small, draw=none},
cycle list name=my colors,
]
\nextgroupplot
\addplot+[const plot, mark=none] table[x=EpTO-50-1sec-x, y=EpTO-50-1sec-y, col sep=comma] {csv-data/plot-local-deltas-cdf.csv};
\addplot+[const plot, mark=none] table[x=JGroups-50-1sec-x, y=JGroups-50-1sec-y, col sep=comma] {csv-data/plot-local-deltas-cdf.csv};
\legend{EpTO, JGroups}
%
\nextgroupplot
\addplot+[const plot, mark=none] table[x=EpTO-50-2sec-x, y=EpTO-50-2sec-y, col sep=comma] {csv-data/plot-local-deltas-cdf.csv};
\addplot+[const plot, mark=none] table[x=JGroups-50-2sec-x, y=JGroups-50-2sec-y, col sep=comma] {csv-data/plot-local-deltas-cdf.csv};
%
\nextgroupplot
\addplot+[const plot, mark=none] table[x=EpTO-100-1sec-x, y=EpTO-100-1sec-y, col sep=comma] {csv-data/plot-local-deltas-cdf.csv};
\addplot+[const plot, mark=none] table[x=JGroups-100-1sec-x, y=JGroups-100-1sec-y, col sep=comma] {csv-data/plot-local-deltas-cdf.csv};
\end{groupplot}
%
\node[anchor=south] at (group c1r1.north) {50-1sec};
\node[anchor=south] at (group c2r1.north) {50-2sec};
\node[anchor=south] at (group c3r1.north) {100-1sec};
\end{tikzpicture}