### 2016-12-01 report/figures/local-times-churn.tex

\begin{tikzpicture}
\begin{groupplot}[
group style={
group size=2 by 1,
vertical sep=0pt,
horizontal sep=6mm,
xlabels at=edge bottom,
ylabels at=edge left,
xticklabels at=edge bottom,
yticklabels at=edge left,
},
every axis plot/.append style={very thick},
height=5cm, width=\linewidth / 1.75,
grid=major,
grid style={dashed,gray!30},
ymax=1.05,
ymin=0,
xmin=1185,
xmax=1202.5,
xlabel={Time $\left[\si{\second}\right]$},
x tick label style={/pgf/number format/fixed},
ytick={0,0.2,0.4,0.6,0.8,1},
yticklabels={0,0.2,0.4,0.6,0.8,1},
ylabel=Percentiles,
legend columns=3,
legend cell align=left,
legend style={at={(1.1,1.3)},anchor=north, font=\small, draw=none},
cycle list name=my colors,
]
\nextgroupplot
\addplot+[const plot, mark=none] table[x=EpTO-suspend-x, y=EpTO-suspend-y, col sep=comma] {csv-data/plot-local-time-cdf.csv};
\addplot+[const plot, mark=none] table[x=JGroups-suspend-coord-x, y=JGroups-suspend-coord-y, col sep=comma] {csv-data/plot-local-time-cdf.csv};
\addplot+[const plot, mark=nonee] table[x=JGroups-suspend-nocoord-x, y=JGroups-suspend-nocoord-y, col sep=comma] {csv-data/plot-local-time-cdf.csv};
\legend{EpTO, JGroups-nocoord, JGroups-coord}
%
\nextgroupplot
\addplot+[const plot, mark=nonek] table[x=EpTO-add-suspend-x, y=EpTO-add-suspend-y, col sep=comma] {csv-data/plot-local-time-cdf.csv};
\addplot+[const plot, mark=none] table[x=JGroups-add-suspend-coord-x, y=JGroups-add-suspend-coord-y, col sep=comma] {csv-data/plot-local-time-cdf.csv};
\addplot+[const plot, mark=none] table[x=JGroups-add-suspend-nocoord-x, y=JGroups-add-suspend-nocoord-y, col sep=comma] {csv-data/plot-local-time-cdf.csv};
\end{groupplot}
%
\node[anchor=south] at (group c1r1.north) {1 kill/minute};
\node[anchor=south] at (group c2r1.north) {1 \{kill,add\}/minute};
\end{tikzpicture}