### 2016-12-01 report/figures/total-bytes-churn.tex

\begin{tikzpicture}
\usetikzlibrary{plotmarks}
\pgfplotsset{width=\linewidth, height=4.7cm}
\begin{groupplot}[
group style={
	group size=2 by 1,
	vertical sep=0pt,
	horizontal sep=2mm,
	xlabels at=edge bottom,
	ylabels at=edge left,
	xticklabels at=edge bottom,
	yticklabels at=edge left,
},
ymin=0,
ymax=35,
width=\linewidth / 1.75,
enlarge x limits=0.3,
ybar=0,
/pgf/bar width=3mm,
/pgfplots/area legend,
nodes near coords,
legend style={
	anchor=north west,
	at={(0.6,0.97)},
	cells={anchor=west},
	draw=none,
},
every node near coord/.append style={
	rotate=90,
	anchor=north,
	font=\tiny,
	xshift=3mm,
	yshift=0.3mm,
},
ymajorgrids,
xtick=data,
xlabel=Protocol,
xticklabel style={text height=1.5ex, rotate=30}, 
ylabel={Total $\left[\si{\giga\byte}\right]$},
symbolic x coords={EpTO, JGroups-nocoord, JGroups-coord},
]
\nextgroupplot
% 100-1sec 1kill/min
\addplot+[mark=none, pattern=north east lines,pattern color=blue, error bars/.cd,y dir=both, y explicit] coordinates {
	% Receive
	(EpTO, 21.3437665258) +- (0, 0.302297622101786)
	(JGroups-nocoord, 1.3338886813) +- (0, 0.00853400183682598)
	(JGroups-coord, 1.3440551501) +- (0, 0.0164341768113215)
};
\addplot+[mark=none, pattern=crosshatch dots,pattern color=red, error bars/.cd,y dir=both, y explicit] coordinates {
	% Send
	(EpTO, 21.4504884001) +- (0, 0.0157182231373203)
	(JGroups-nocoord, 1.285075604) +- (0, 0.00872334968295833)
	(JGroups-coord, 1.2945710642) +- (0, 0.0158581561021704)
};
\legend{receive, send}
\nextgroupplot
% 100-1sec 1kill, 1add/min
\addplot+[mark=none, pattern=north east lines,pattern color=blue, error bars/.cd,y dir=both, y explicit] coordinates {
	% Receive
	(EpTO, 27.0797668262) +- (0, 0.300652054138594)
	(JGroups-nocoord, 1.6071496365) +- (0, 0.00972747186106245)
	(JGroups-coord, 1.6039433255) +- (0, 0.0084962148465832)
};
\addplot+[mark=none, pattern=crosshatch dots,pattern color=red, error bars/.cd,y dir=both, y explicit] coordinates {
	% Send
	(EpTO, 26.8856375946) +- (0, 0.0157182231373203)
	(JGroups-nocoord, 1.5459290852) +- (0, 0.00937891633596378)
	(JGroups-coord, 1.5426003209) +- (0, 0.00829044483237941)
};
\end{groupplot}
%
\node[anchor=south] at (group c1r1.north) {1 kill/minute};
\node[anchor=south] at (group c2r1.north) {1 kill/minute, 1 add/minute};
\end{tikzpicture}