### 2016-12-01 report/figures/total-bytes.tex

\begin{tikzpicture}
\usetikzlibrary{plotmarks}
\pgfplotsset{width=\linewidth, height=4.7cm}
\begin{groupplot}[
group style={
	group size=3 by 1,
	vertical sep=0pt,
	horizontal sep=2mm,
	xlabels at=edge bottom,
	ylabels at=edge left,
	xticklabels at=edge bottom,
	yticklabels at=edge left,
},
ymin=0,
ymax=35,
width=\linewidth / 2.5,
enlarge x limits=0.3,
ybar=0,
/pgf/bar width=3mm,
/pgfplots/area legend,
nodes near coords,
legend style={
	anchor=north west,
	at={(0.3,0.97)},
	cells={anchor=west},
	draw=none,
},
every node near coord/.append style={
	rotate=90,
	anchor=north,
	font=\tiny,
	xshift=3mm,
	yshift=0.3mm,
},
ymajorgrids,
xtick=data,
xlabel=Protocol,
ylabel={Total $\left[\si{\giga\byte}\right]$},
symbolic x coords={EpTO, JGroups},
]
\nextgroupplot
% 50-1sec
\addplot+[mark=none, pattern=north east lines,pattern color=blue, error bars/.cd,y dir=both, y explicit] coordinates {
	% Receive
	(EpTO, 10.8956669416) +- (0, 0.112678096321221)
	(JGroups, 0.6420599937) +- (0, 0.0164341768113215)
};
\addplot+[mark=none, pattern=crosshatch dots,pattern color=red, error bars/.cd,y dir=both, y explicit] coordinates {
	% Send
	(EpTO, 10.8952655703) +- (0, 0.112677849128705)
	(JGroups, 0.628793349) +- (0, 0.0164560780074577)
};
\legend{receive, send}
\nextgroupplot
% 50-2sec
\addplot+[mark=none, pattern=north east lines,pattern color=blue, error bars/.cd,y dir=both, y explicit] coordinates {
	% Receive
	(EpTO, 22.1900560357) +- (0, 0.513548491076014)
	(JGroups, 1.1888473068) +- (0, 0.0236261886904278)
};
\addplot+[mark=none, pattern=crosshatch dots,pattern color=red, error bars/.cd,y dir=both, y explicit] coordinates {
	% Send
	(EpTO, 22.1896583376) +- (0, 0.513548672987696)
	(JGroups, 1.1756330318) +- (0, 0.0236412207168411)
};
\nextgroupplot
% 100-1sec
\addplot+[mark=none, pattern=north east lines,pattern color=blue, error bars/.cd,y dir=both, y explicit] coordinates {
	% Receive
	(EpTO, 25.9895900054) +- (0, 0.285368750128887)
	(JGroups, 1.6842673321) +- (0, 0.014352534074787)
};
\addplot+[mark=none, pattern=crosshatch dots,pattern color=red, error bars/.cd,y dir=both, y explicit] coordinates {
	% Send
	(EpTO, 25.9864516777) +- (0, 0.285365521649361)
	(JGroups, 1.6291726389) +- (0, 0.0124299318567327)
};
\end{groupplot}
%
\node[anchor=south] at (group c1r1.north) {50-1sec};
\node[anchor=south] at (group c2r1.north) {50-2sec};
\node[anchor=south] at (group c3r1.north) {100-1sec};
\end{tikzpicture}