### 2016-12-01 report/figures/total-events-sent-churn.tex

\begin{tikzpicture}
\pgfplotsset{width=\linewidth, height=4.7cm}
\begin{groupplot}[
group style={
	group size=2 by 1,
	vertical sep=0pt,
	horizontal sep=2mm,
	xlabels at=edge bottom,
	ylabels at=edge left,
	xticklabels at=edge bottom,
	yticklabels at=edge left,
},
ymin=53000,
ymax=65000,
width=\linewidth / 1.75,
enlarge x limits=0.3,
ybar=0,
/pgf/bar width=3mm,
/pgfplots/area legend,
nodes near coords,
legend style={
	anchor=north west,
	at={(0.6,0.97)},
	cells={anchor=west},
	draw=none,
},
every node near coord/.append style={
	rotate=90,
	anchor=north,
	font=\tiny,
	xshift=5mm,
	yshift=1mm,
},
ymajorgrids,
xtick=data,
xlabel=Protocol,
xticklabel style={text height=1.5ex, rotate=30}, 
scaled ticks=false, tick label style={/pgf/number format/fixed},
ylabel={Events sent},
symbolic x coords={EpTO, JGroups-nocoord, JGroups-coord},
]
\nextgroupplot
% 100-1sec 1kill/min
\addplot[mark=none, pattern=north east lines, error bars/.cd,y dir=both, y explicit] coordinates {
	% Receive
	(EpTO, 54169.6) +- (0, 158.689214084218)
	(JGroups-nocoord, 53981.9) +- (0, 102.155056436554)
	(JGroups-coord, 53953.4) +- (0, 180.331300050151)
};
%
\nextgroupplot
% 100-1sec 1kill, 1add/min
\addplot[mark=none, pattern=north east lines, error bars/.cd,y dir=both, y explicit] coordinates {
	% Receive
	(EpTO, 59853.9) +- (0, 228.339148344445)
	(JGroups-nocoord, 57813.8) +- (0, 128.414087146925)
	(JGroups-coord, 57873.3) +- (0, 150.893671172785)
};
\end{groupplot}
%
\node[anchor=south] at (group c1r1.north) {1 kill/minute};
\node[anchor=south] at (group c2r1.north) {1 \{kill,add\}/minute};
\end{tikzpicture}