### 2016-12-22 report/figures/no-churn/total-bytes.tex

\begin{tikzpicture}
\usetikzlibrary{plotmarks}
\pgfplotsset{width=\linewidth, height=4.7cm}
\begin{groupplot}[
group style={
	group size=3 by 1,
	vertical sep=0pt,
	horizontal sep=2mm,
	xlabels at=edge bottom,
	ylabels at=edge left,
	xticklabels at=edge bottom,
	yticklabels at=edge left,
},
ymin=0,
ymax=35,
width=\linewidth / 2.5,
enlarge x limits=0.3,
ybar=0,
/pgf/bar width=3mm,
/pgfplots/area legend,
nodes near coords,
legend style={
	anchor=north west,
	at={(0.3,0.97)},
	cells={anchor=west},
	draw=none,
},
every node near coord/.append style={
	rotate=90,
	anchor=north,
	font=\tiny,
	xshift=3mm,
	yshift=0.3mm,
},
ymajorgrids,
xtick=data,
xlabel=Protocol,
ylabel={Total $\left[\si{\giga\byte}\right]$},
symbolic x coords={EpTO, JGroups},
]
\nextgroupplot
% 50-1sec
\addplot+[mark=none, pattern=north east lines,pattern color=blue, error bars/.cd,y dir=both, y explicit] coordinates {
	% Receive
	(EpTO, 10.8454953841) +- (0, 0.158556056880199)
	(JGroups, 0.8112389025) +- (0, 0.0275185471028981)
};
\addplot+[mark=none, pattern=crosshatch dots,pattern color=red, error bars/.cd,y dir=both, y explicit] coordinates {
	% Send
	(EpTO, 10.84508943053) +- (0, 0.158553532839913)
	(JGroups, 0.7979795057) +- (0, 0.0275124778152395)
};
\legend{receive, send}
\nextgroupplot
% 50-2sec
\addplot+[mark=none, pattern=north east lines,pattern color=blue, error bars/.cd,y dir=both, y explicit] coordinates {
	% Receive
	(EpTO, 22.3122950339) +- (0, 0.387040306037242)
	(JGroups, 1.481260734) +- (0, 0.0112693613185708)
};
\addplot+[mark=none, pattern=crosshatch dots,pattern color=red, error bars/.cd,y dir=both, y explicit] coordinates {
	% Send
	(EpTO, 22.3118855827) +- (0, 0.387040491969926)
	(JGroups, 1.4680794813) +- (0, 0.0112627440922006)
};
\nextgroupplot
% 100-1sec
\addplot+[mark=none, pattern=north east lines,pattern color=blue, error bars/.cd,y dir=both, y explicit] coordinates {
	% Receive
	(EpTO, 26.0194016553) +- (0, 0.2773178289900919968)
	(JGroups, 2.0044813206) +- (0, 0.0103933483299930001)
};
\addplot+[mark=none, pattern=crosshatch dots,pattern color=red, error bars/.cd,y dir=both, y explicit] coordinates {
	% Send
	(EpTO, 26.0161708501) +- (0, 0.277318407393540023)
	(JGroups, 1.9504773287) +- (0, 0.0104754176417391)
};
\end{groupplot}
%
\node[anchor=south] at (group c1r1.north) {50-1sec};
\node[anchor=south] at (group c2r1.north) {50-2sec};
\node[anchor=south] at (group c3r1.north) {100-1sec};
\end{tikzpicture}