### 2016-12-22 report/figures/synthetic-churn/bandwidth-figure-churn.tex

\begin{tikzpicture}[domain=0:1800]
\usetikzlibrary{plotmarks}
\pgfplotsset{
	height=4cm,
	width=\linewidth / 2.5,
	every axis plot post/.append style={
		solid,
		very thin,
		mark=none
	},
	/pgfplots/area cycle list/.style={/pgfplots/cycle list={%
			{black,fill=black,mark=none},%
			{black,fill=white!25!black,mark=none},%
			{black,fill=white!50!black,mark=none},%
			{black,fill=white!75!black,mark=none},%
			{black,fill=white,mark=none},%
		}
	},
}
\begin{groupplot}[
ymajorgrids,
group style={
	group size=2 by 3,
	vertical sep=8mm,
	horizontal sep=4mm,
	xlabels at=edge bottom,
	ylabels at=edge left,
	yticklabels at=edge left,
},
stack plots=y,area style, enlarge x limits=false, 
ymin=0,
xmin=0,
ylabel={Bandwidth $\left[\SI{}{\mbps}\right]$},
xlabel={Time $\left[\si{\minute}\right]$},
legend columns=5,
legend cell align=left,
legend style={at={(1.25,1.5)},anchor=north, font=\small, draw=none},]
\nextgroupplot[ymax=4,ytick={0,1,2,3,4},
yticklabels={0,1,2,3,4},]
\addplot table[x=time,y=EpTO-suspend-0.000000, col sep=comma]{\tableaverage} \closedcycle;
\addplot table[x=time,y=EpTO-suspend-0.250000, col sep=comma]{\tableaverage} \closedcycle;
\addplot table[x=time,y=EpTO-suspend-0.500000, col sep=comma]{\tableaverage} \closedcycle;
\addplot table[x=time,y=EpTO-suspend-0.750000, col sep=comma]{\tableaverage} \closedcycle;
\addplot table[x=time,y=EpTO-suspend-1.000000, col sep=comma]{\tableaverage} \closedcycle;
\legend{0, 0.25, 0.5, 0.75, 1}
%
\nextgroupplot[ymax=4,ytick={0,1,2,3,4},]
\addplot table[x=time,y=EpTO-add-suspend-0.000000, col sep=comma]{\tableaverage} \closedcycle;
\addplot table[x=time,y=EpTO-add-suspend-0.250000, col sep=comma]{\tableaverage} \closedcycle;
\addplot table[x=time,y=EpTO-add-suspend-0.500000, col sep=comma]{\tableaverage} \closedcycle;
\addplot table[x=time,y=EpTO-add-suspend-0.750000, col sep=comma]{\tableaverage} \closedcycle;
\addplot table[x=time,y=EpTO-add-suspend-1.000000, col sep=comma]{\tableaverage} \closedcycle;
%
\nextgroupplot[ymax=20,ytick={0,5,10,15,20},
yticklabels={0,5,10,15,20},]
\addplot table[x=time,y=JGroups-suspend-coord-0.000000, col sep=comma]{\tableaverage} \closedcycle;
\addplot table[x=time,y=JGroups-suspend-coord-0.250000, col sep=comma]{\tableaverage} \closedcycle;
\addplot table[x=time,y=JGroups-suspend-coord-0.500000, col sep=comma]{\tableaverage} \closedcycle;
\addplot table[x=time,y=JGroups-suspend-coord-0.750000, col sep=comma]{\tableaverage} \closedcycle;
\addplot table[x=time,y=JGroups-suspend-coord-1.000000, col sep=comma]{\tableaverage} \closedcycle;
%
\nextgroupplot[ymax=20,ytick={0,5,10,15,20},]
\addplot table[x=time,y=JGroups-add-suspend-coord-0.000000, col sep=comma]{\tableaverage} \closedcycle;
\addplot table[x=time,y=JGroups-add-suspend-coord-0.250000, col sep=comma]{\tableaverage} \closedcycle;
\addplot table[x=time,y=JGroups-add-suspend-coord-0.500000, col sep=comma]{\tableaverage} \closedcycle;
\addplot table[x=time,y=JGroups-add-suspend-coord-0.750000, col sep=comma]{\tableaverage} \closedcycle;
\addplot table[x=time,y=JGroups-add-suspend-coord-1.000000, col sep=comma]{\tableaverage} \closedcycle;
%
\nextgroupplot[ymax=2,ytick={0,0.5,1,1.5,2},
yticklabels={0,0.5,1,1.5,2},]
\addplot table[x=time,y=JGroups-suspend-nocoord-0.000000, col sep=comma]{\tableaverage} \closedcycle;
\addplot table[x=time,y=JGroups-suspend-nocoord-0.250000, col sep=comma]{\tableaverage} \closedcycle;
\addplot table[x=time,y=JGroups-suspend-nocoord-0.500000, col sep=comma]{\tableaverage} \closedcycle;
\addplot table[x=time,y=JGroups-suspend-nocoord-0.750000, col sep=comma]{\tableaverage} \closedcycle;
\addplot table[x=time,y=JGroups-suspend-nocoord-1.000000, col sep=comma]{\tableaverage} \closedcycle;
%
\nextgroupplot[ymax=2,ytick={0,0.5,1,1.5,2},]
\addplot table[x=time,y=JGroups-add-suspend-nocoord-0.000000, col sep=comma]{\tableaverage} \closedcycle;
\addplot table[x=time,y=JGroups-add-suspend-nocoord-0.250000, col sep=comma]{\tableaverage} \closedcycle;
\addplot table[x=time,y=JGroups-add-suspend-nocoord-0.500000, col sep=comma]{\tableaverage} \closedcycle;
\addplot table[x=time,y=JGroups-add-suspend-nocoord-0.750000, col sep=comma]{\tableaverage} \closedcycle;
\addplot table[x=time,y=JGroups-add-suspend-nocoord-1.000000, col sep=comma]{\tableaverage} \closedcycle;
%
\end{groupplot}
\node[anchor=south] at (group c1r1.north) {1 kill};
\node[anchor=south] at (group c2r1.north) {1 \{kill,add\}/minute};
\node[anchor=south, rotate=-90] at (group c2r1.east){EpTO};
\node[anchor=south, rotate=-90] at (group c2r2.east){JGroups coord};
\node[anchor=south, rotate=-90] at (group c2r3.east){JGroups nocoord};
\end{tikzpicture}