### 2016-12-22 report/main.tex

% !TeX program = lualatex
\documentclass[10pt,conference,a4paper]{IEEEtran}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage[rm={oldstyle=false}]{cfr-lm}
\usepackage[australian,american]{babel}

\usepackage[backend=biber,style=ieee,bibencoding=utf8,sorting=none,doi=false,isbn=false,url=false,date=short]{biblatex}
\usepackage{amssymb}
\usepackage[cmex10]{amsmath}
\usepackage{csquotes}
\usepackage[nolist]{acronym}
\usepackage{float}
\usepackage{todonotes}
\usepackage{pgfplots}
\usepackage{pgf}
\usepackage[binary-units]{siunitx}
\usepackage{pgfplotstable}
\usepackage{booktabs}
\usepackage{multirow}
\usepackage{lipsum}
\usepackage[hidelinks]{hyperref}
\usepackage{xspace}
\usepackage[caption=false]{subfig}

\usetikzlibrary{patterns}

\usepgfplotslibrary{external, groupplots, units}
\tikzexternalize
\tikzsetexternalprefix{generated-figures/}

\sisetup{
	round-mode          = places,
	round-precision     = 2,
}
\DeclareSIUnit\mbps{MB/s}

\pgfplotscreateplotcyclelist{my colors}{%
solid, color=black\\%
densely dotted, color=blue\\%
densely dashed, color=red\\%
}
\makeatother
%% Environment for comments: Set the boolean to false to produce a comment-free version.
\newboolean{showcomments}
\setboolean{showcomments}{true}
\ifthenelse{\boolean{showcomments}}
{ \newcommand{\mynote}[3]{
		\fbox{\bfseries\sffamily\scriptsize#1}
		{\small$\blacktriangleright$\textsf{\textit{\color{#3}{#2}}}$\blacktriangleleft$}}}
{ \newcommand{\mynote}[3]{}}
% One command per author:
\newcommand{\jt}[1]{\mynote{Jocelyn}{#1}{blue}}

\newcommand{\epto}{\textsc{EpTO}\xspace}
\newcommand{\jgroups}{\textsc{JGroups}\xspace}

\newif\ifjmcs
% Change false to true to have the JMCS cover page
\jmcstrue

\pgfplotsset{
	compat=1.14,
	tick label style={font=\small},
	label style={font=\small},
	legend style={font=\small},
	/pgfplots/short line legend/.style={
		legend image code/.code={
			\draw[mark repeat=2,mark phase=2,##1]
			plot coordinates {
				(0cm,0cm)
				(0.2cm,0cm)
				(0.4cm,0cm)
			};%
		}
	},
	/pgfplots/xtick pos=left,
}

\def\ieeetitle{EpTO TO DEFINE}
\def\ieeeauthor{X,X,X}

\def\thesistitle{EpTO TO DEFINE}
\def\thesissubtitle{X}
\def\thesisauthor{Jocelyn Thode}

\author{\IEEEauthorblockN{\ieeeauthor}
	\IEEEauthorblockA{University of Fribourg, Switzerland\\
		\href{mailto:first.last@unifr.ch}{first.last@unifr.ch}}
}
\title{\ieeetitle}

\addbibresource{references.bib}
\renewcommand*{\bibfont}{\small}

\hypersetup{
	pdftitle=\thesistitle,
	pdfauthor=\thesisauthor
}

% \noautocite{*}

\begin{document}

\ifjmcs
\begin{titlepage}
	\begin{otherlanguage}{australian}
		\begin{center}
			\begin{figure}[t]
				\center{\includegraphics[scale=0.2]{logos/MSc_quer.png}}
				\vspace{0.4in}
			\end{figure}

			{\bfseries\Huge \thesistitle \par
				\Large \vspace{0.1in} \thesissubtitle \par}

			\vspace{0.3in}
			\LARGE{\textbf{Master Thesis} \\}
			\vspace{0.4in}

			{\Large \thesisauthor}

			\vspace{0.3in}
			{\Large Université de Fribourg \par}
			\vfill
			{\Large \today \par}

			\vspace{0.9in}

			% === Logos ==============================================
			\begin{figure}[htp]
				\centering
				\subfloat{\includegraphics[scale=0.60]{logos/UNI_Bern.pdf}}\hfill
				\subfloat{\includegraphics[scale=0.54]{logos/UNI_Neuchatel.pdf}}\hfill
				\subfloat{\includegraphics[scale=0.81]{logos/UNI_Fribourg.pdf}}
			\end{figure}
			% === // Logos ===========================================

		\end{center}
	\end{otherlanguage}
\end{titlepage}

\pagebreak
\fi

%	\maketitle

\input{abstract.tex}
\input{introduction.tex}
\input{ordering.tex}
\input{definitions.tex}
\input{framework.tex}
\input{evaluation.tex}
\input{future-work.tex}
\input{conclusion.tex}

\printbibliography
\end{document}