### 2016-12-23 report/conclusion.tex
Continue work on thesis

\section{Conclusion}
\label{sec:conclusion}
\jt{Do later}