### 2016-12-23 report/figures/architecture.tex

	\begin{tikzpicture}[transform shape,scale=0.6]
\node[draw, text width=2.1cm, text centered] (bench) at (2.1, 2.6) {Benchmarks runner};
\draw (0.8,3.4) node[below right,red,scale=0.8] {\textit{Python script}};
\draw[dotted,red,thick] (0.8,2) rectangle (3.4,3.4);
%
\draw[loosely dashed,black] (3.7,4.3) rectangle (14.8,-0.8);
\draw (3.7,4.3) node[below right,black,scale=0.8] {\textit{Swarm overlay:} \texttt{epto containers communicate between each other using UDP}};
%
\node[draw] (tck) at (12.4, 2.5) {Tracker};
\draw (11.6,3.25) node[below right,red,scale=0.8] {\textit{Python script}};
\draw (11.5,3.65) node[below right,blue,scale=0.8] {\textit{Docker:} \texttt{epto-tracker:}};
\draw[loosely dashed,blue] (11.5,1.9) rectangle (14.5,3.65);
\draw[dotted,red,thick] (11.6,2) rectangle (14.4,3.25);
%
\node[draw] (app1) at (7.1, 2.5) {EpTOTester 1};
\draw (5.7,3.65) node[below right,blue,scale=0.8] {\textit{Docker:} \texttt{epto}};
\draw (5.8,3.25) node[below right,red,scale=0.8] {\textit{Java program}};
\draw[loosely dashed,blue] (5.7,2) rectangle (8.5,3.65);
\draw[dotted,red,thick] (5.8,2.1) rectangle (8.4,3.25);
%
\node[draw] (app2) at (5.4, 0) {EpTOTester 2};
\draw (4,1.15) node[below right,blue,scale=0.8] {\textit{Docker:} \texttt{epto}};
\draw (4.1,.75) node[below right,red,scale=0.8] {\textit{Java program}};
\draw[loosely dashed,blue] (4,-0.5) rectangle (6.8,1.15);
\draw[dotted,red,thick] (4.1,-0.4) rectangle (6.7,.75);
%
\node[draw] (appn) at (9.4, 0) {EpTOTester N};
\draw (8,1.15) node[below right,blue,scale=0.8] {\textit{Docker:} \texttt{epto}};
\draw (8.1,.75) node[below right,red,scale=0.8] {\textit{Java program}};
\draw[loosely dashed,blue] (8,-0.5) rectangle (10.8,1.15);
\draw[dotted,red,thick] (8.1,-0.4) rectangle (10.7,.75);
%
\draw[<->,thick,>=stealth] (app1.east) to[out=5,in=180] node[sloped,midway,scale=0.6,above]{HTTP} (tck.west);
\draw[<->,thick,>=stealth] (app2.east) to[out=50,in=180] node[sloped,midway,scale=0.6,above]{HTTP} (tck.west);
\draw[<->,thick,>=stealth] (appn.east) to[out=5,in=180] node[sloped,midway,scale=0.6,above]{HTTP} (tck.west);
%
\draw[<->,thick,>=stealth] (bench.east) to[out=5,in=180] node[sloped,midway,scale=0.6,above]{Start/Stop} (app1.west);
\draw[<->,thick,>=stealth] (bench.east) to[out=5,in=180] node[sloped,midway,scale=0.6,above]{Start/Stop} (app2.west);
\draw[<->,thick,>=stealth] (bench.east) to[out=5,in=180] node[sloped,midway,scale=0.6,above]{Start/Stop} (appn.west);
\end{tikzpicture}