### 2016-12-23 report/figures/real-churn/local-delta-real-churn.tex

\begin{tikzpicture}
\begin{axis}[
every axis plot/.append style={very thick},
height=5cm,
grid=major,
grid style={dashed,gray!30},
ymax=1.05,
ymin=0,
scaled x ticks = false,
x tick label style={/pgf/number format/fixed},
xlabel={Time $\left[\si{\second}\right]$},
ytick={0,0.2,0.4,0.6,0.8,1},
yticklabels={0,0.2,0.4,0.6,0.8,1},
ylabel=Percentiles,
legend columns=3,
legend cell align=left,
legend style={at={(0.5,1.2)},anchor=north, font=\small, draw=none},
cycle list name=my colors,
]
\addplot+[const plot, mark=none] table[x=EpTO-real-trace-x, y=EpTO-real-trace-y, col sep=comma] {\tablelocaldeltas};
\addplot+[const plot, mark=none] table[x=JGroups-coord-x, y=JGroups-coord-y, col sep=comma] {\tablelocaldeltas};
\addplot+[const plot, mark=nonee] table[x=JGroups-nocoord-x, y=JGroups-nocoord-y, col sep=comma] {\tablelocaldeltas};
\legend{EpTO, JGroups-coord, JGroups-nocoord}
\end{axis}
\end{tikzpicture}