### 2016-12-23 report/future-work.tex

\section{Future Work}
\label{sec:future-work}
\jt{Speak about running test on a bigger infrastructure, because the small ammount of nodes does not let us see the full potential of \epto. Then speak about the ongoing work of Push-Pull}