### 2016-12-27 report/figures/real-churn/bandwidth-figure-real-churn.tex

\begin{tikzpicture}
\usetikzlibrary{plotmarks}
\pgfplotsset{
	height=4cm,
	width=\linewidth,
	every axis plot post/.append style={
		solid,
		very thin,
		mark=none
	},
	/pgfplots/area cycle list/.style={/pgfplots/cycle list={%
			{black,fill=black,mark=none},%
			{black,fill=white!25!black,mark=none},%
			{black,fill=white!50!black,mark=none},%
			{black,fill=white!75!black,mark=none},%
			{black,fill=white,mark=none},%
		}
	},
}
\begin{groupplot}[
ymajorgrids,
group style={
	group size=1 by 4,
	vertical sep=6mm,
	horizontal sep=0mm,
	xlabels at=edge bottom,
	xticklabels at=edge bottom,
	ylabels at=edge left,
},
stack plots=y,area style, enlarge x limits=false, 
ymin=0,
xmin=0,
xmax=61,
ylabel={Bandwidth $\left[\SI{}{\mbps}\right]$},
xlabel={Time $\left[\si{\minute}\right]$},
legend columns=5,
legend cell align=left,
legend style={at={(0.6,1.9)},anchor=north, font=\small, draw=none},]
\nextgroupplot[height=2.3cm, 
ylabel={Nodes}, 
ymin=75,ymax=120,
ytick={80,100,120}, 
yticklabels={80,100,120},
xmajorgrids,
yminorgrids,
tick label style={font=\footnotesize},
label style={font=\tiny}]
\addplot[const plot, color=blue, mark=none,set layers,
axis lines = middle,] table [x=minute, y=size, col sep=comma] {csv-data/real-trace.csv};
\nextgroupplot[ymax=4,ytick={0,1,2,3,4},
yticklabels={0,1,2,3,4}]
\addplot table[x=time,y=EpTO-real-trace-0.000000, col sep=comma]{\tableaveragereal} \closedcycle;
\addplot table[x=time,y=EpTO-real-trace-0.250000, col sep=comma]{\tableaveragereal} \closedcycle;
\addplot table[x=time,y=EpTO-real-trace-0.500000, col sep=comma]{\tableaveragereal} \closedcycle;
\addplot table[x=time,y=EpTO-real-trace-0.750000, col sep=comma]{\tableaveragereal} \closedcycle;
\addplot table[x=time,y=EpTO-real-trace-1.000000, col sep=comma]{\tableaveragereal} \closedcycle;
\legend{0, 0.25, 0.5, 0.75, 1}
%
\nextgroupplot%[ymax=4,ytick={0,1,2,3,4},]
\addplot table[x=time,y=JGroups-coord-0.000000, col sep=comma]{\tableaveragereal} \closedcycle;
\addplot table[x=time,y=JGroups-coord-0.250000, col sep=comma]{\tableaveragereal} \closedcycle;
\addplot table[x=time,y=JGroups-coord-0.500000, col sep=comma]{\tableaveragereal} \closedcycle;
\addplot table[x=time,y=JGroups-coord-0.750000, col sep=comma]{\tableaveragereal} \closedcycle;
\addplot table[x=time,y=JGroups-coord-1.000000, col sep=comma]{\tableaveragereal} \closedcycle;
%
\nextgroupplot[ymax=4,ytick={0,1,2,3,4}]
\addplot table[x=time, y=JGroups-nocoord-0.000000, col sep=comma]{\tableaveragereal} \closedcycle;
\addplot table[x=time,y=JGroups-nocoord-0.250000, col sep=comma]{\tableaveragereal} \closedcycle;
\addplot table[x=time,y=JGroups-nocoord-0.500000, col sep=comma]{\tableaveragereal} \closedcycle;
\addplot table[x=time,y=JGroups-nocoord-0.750000, col sep=comma]{\tableaveragereal} \closedcycle;
\addplot table[x=time,y=JGroups-nocoord-1.000000, col sep=comma]{\tableaveragereal} \closedcycle;
\end{groupplot}
%
\node[anchor=south, rotate=-90] at (group c1r2.east) {\epto};
\node[anchor=south, rotate=-90] at (group c1r3.east) {\jgroups-coord};
\node[anchor=south, rotate=-90] at (group c1r4.east) {\jgroups-nocoord};
\end{tikzpicture}