### 2016-12-28 report/future-work.tex

\section{Future Work}
\label{sec:future-work}
\jt{Do we have Push-Pull references?}
Future work should focus on testing \epto on a bigger cluster, preferably at a point where \jgroups can no longer function, as \epto is designed to work on very large number of peers. As it is, \epto does indeed work as intended in our infrastructure and under churn, but a larger test would showcase \epto's ability to handle large peer numbers.
\par
Future work also includes working on an alternate \epto dissemination algorithm. This algorithm would use a Push-Pull approach to further reduce the needed number of balls in the system and thus reduce the network load.