### 2016-12-30 report/abstract.tex
Fix more stuff pointed by Hugues

\begin{abstract}
  \epto  \autocite{matos2015epto} is one of the recently introduced total order algorithms for large-scale distributed systems, and provides total order and scalability at the same time. In this paper, we verify this claim by implementing EpTO and evaluate its reliability in real-world conditions by comparing it to a deterministic total order algorithm named \jgroups. We first compare them in a perfect environment assessing the scalability in terms of events throughput and peer numbers. We then compare them using a synthetic churn trace and a real one using the same metrics. Our results show \epto is performing has expected, although further testing on a bigger cluster is required to evaluate how it performs when \jgroups no longer does.
  \hm{General comment: if possible, maybe we should rent resources on Amazon to test with larger networks.}
\end{abstract}