### 2017-01-12 report/abstract.tex
Take into account all remaining comments from Miguel

\begin{abstract}
  \epto  \autocite{matos2015epto} is one of the recently introduced total order algorithms for large-scale distributed systems, and provides total order and scalability at the same time. In this paper, we verify this claim by designing an architecture to facilitate the evaluation of large scale distributed systems, implement \epto and evaluate its reliability in real-world conditions by comparing it to a deterministic total order algorithm named \jgroups. We first compare them in a perfect environment assessing the scalability in terms of events throughput and peer numbers. We then compare them using a synthetic churn trace and a real one using the same metrics. Our results show \epto is performing has expected, although further testing on a bigger cluster is required to evaluate how it performs when \jgroups no longer does.
  \hm{General comment: if possible, maybe we should rent resources on Amazon to test with larger networks.}
\end{abstract}