### 2017-02-06 report/figures/no-churn/total-events-sent.tex

\begin{tikzpicture}
\usetikzlibrary{plotmarks}
\pgfplotsset{width=\linewidth, height=4.7cm}
\begin{groupplot}[
group style={
	group size=3 by 1,
	vertical sep=0pt,
	horizontal sep=2mm,
	xlabels at=edge bottom,
	ylabels at=edge left,
	xticklabels at=edge bottom,
	yticklabels at=edge left,
},
ymin=40000,
ymax=155000,
width=\linewidth / 2.5,
enlarge x limits=0.3,
ybar=0,
/pgf/bar width=3mm,
/pgfplots/area legend,
nodes near coords,
legend style={
	anchor=north west,
	at={(0.3,0.97)},
	cells={anchor=west},
	draw=none,
	fill=none,
},
every node near coord/.append style={
	rotate=90,
	anchor=north,
	font=\tiny,
	xshift=5mm,
	yshift=1mm,
},
ymajorgrids,
xtick=data,
xlabel=Protocol,
ylabel={Events sent},
scaled ticks=false, tick label style={/pgf/number format/fixed},
symbolic x coords={EpTO, JGroups},
]
\nextgroupplot
% 50-1sec
\addplot[mark=none, pattern=north east lines, error bars/.cd,y dir=both, y explicit] coordinates {
	% Receive
	(EpTO, 59993.8) +- (0, 3.32665998663345)
	(JGroups, 59961.9) +- (0, 10.9285558667798)
};
%
\nextgroupplot
% 50-2sec
\addplot[mark=none, pattern=north east lines, error bars/.cd,y dir=both, y explicit] coordinates {
	% Receive
	(EpTO, 119898.2) +- (0, 9.75021367287237)
	(JGroups, 119885.7) +- (0, 5.01220732035492)
};
\nextgroupplot
% 100-1sec
\addplot[mark=none, pattern=north east lines, error bars/.cd,y dir=both, y explicit] coordinates {
	% Receive
	(EpTO, 59913) +- (0, 164.332319131421)
	(JGroups, 60023.1) +- (0, 287.150928181603)
};
\end{groupplot}
%
\node[anchor=south] at (group c1r1.north) {50-1sec};
\node[anchor=south] at (group c2r1.north) {50-2sec};
\node[anchor=south] at (group c3r1.north) {100-1sec};
\end{tikzpicture}