### 2017-02-06 report/figures/real-churn/total-bytes-real-churn.tex

\begin{tikzpicture}
\usetikzlibrary{plotmarks}
\pgfplotsset{width=\linewidth, height=4.7cm}
\begin{axis}[
ymin=0,
ymax=85,
enlarge x limits=0.3,
ybar=0,
/pgf/bar width=3mm,
/pgfplots/area legend,
nodes near coords,
legend style={
	anchor=north west,
	at={(0.6,0.97)},
	cells={anchor=west},
	draw=none,
	fill=none,
},
every node near coord/.append style={
	rotate=90,
	anchor=north,
	font=\tiny,
	xshift=3mm,
	yshift=0.3mm,
},
ymajorgrids,
xtick=data,
xlabel=Protocol,
xticklabel style={text height=1.5ex, rotate=30}, 
ylabel={Total $\left[\si{\giga\byte}\right]$},
symbolic x coords={EpTO, JGroups-nocoord, JGroups-coord},
]
% 100-1sec 1kill/min
\addplot+[mark=none, pattern=north east lines,pattern color=blue, error bars/.cd,y dir=both, y explicit] coordinates {
	% Receive
	(EpTO, 81.4132665062) +- (0, 1.0822819001435)
	(JGroups-nocoord, 5.6382174422) +- (0, 0.058029870294168)
	(JGroups-coord, 5.6118450728) +- (0, 0.089291024953546)
};
\addplot+[mark=none, pattern=crosshatch dots,pattern color=red, error bars/.cd,y dir=both, y explicit] coordinates {
	% Send
	(EpTO, 82.6741897528) +- (0, 1.08368411575119)
	(JGroups-nocoord, 5.4807749566) +- (0, 0.056977818312573)
	(JGroups-coord, 5.4548362802) +- (0, 0.086273738554255)
};
\legend{receive, send}
\end{axis}
\end{tikzpicture}