### 2017-02-06 report/figures/real-churn/total-events-sent-real-churn.tex

\begin{tikzpicture}
\pgfplotsset{width=\linewidth, height=4.7cm}
\begin{axis}[
ymin=160000,
ymax=170000,
enlarge x limits=0.3,
ybar=0,
/pgf/bar width=3mm,
/pgfplots/area legend,
nodes near coords,
legend style={
	anchor=north west,
	at={(0.6,0.97)},
	cells={anchor=west},
	draw=none,
	fill=none
},
every node near coord/.append style={
	rotate=90,
	anchor=north,
	font=\tiny,
	xshift=5mm,
	yshift=1mm,
},
ymajorgrids,
xtick=data,
xlabel=Protocol,
xticklabel style={text height=1.5ex, rotate=30}, 
scaled ticks=false, tick label style={/pgf/number format/fixed},
ylabel={Events sent},
symbolic x coords={EpTO, JGroups-nocoord, JGroups-coord},
]
% 100-1sec real trace
\addplot[mark=none, pattern=north east lines, error bars/.cd,y dir=both, y explicit] coordinates {
	% Events sent
	(EpTO, 165844.2) +- (0, 210.222976860287)
	(JGroups-nocoord, 166585.8) +- (0, 824.998303028558)
	(JGroups-coord, 166183) +- (0, 1368.95507596122)
};
\end{axis}
\end{tikzpicture}