### 2017-02-07 report/abstract.tex
general pass.

\begin{abstract}
  \epto is one of the recently introduced total order algorithms for large-scale distributed systems, and provides total order and scalability at the same time. In this paper, we verify this claim by designing a framework to facilitate the evaluation of large scale distributed systems, implement \epto and evaluate its reliability in real-world conditions by comparing it to a deterministic total order algorithm provided by the \jgroups toolkit. We first compare them in a perfect environment assessing the scalability in terms of events throughput and number of peers. We then compare them using a synthetic churn trace and a real one using the same metrics. 
Our results show \epto is performing as expected, even though in the system sizes we tested \jgroups is still more efficient in a stable environment.
By extrapolating our results for larger system sizes we can clearly see that \jgroups starts to degrade with the system size. 
Still, further testing on a bigger cluster is required to evaluate how \epto performs when \jgroups is no longer able to keep up with scale.
  \hm{General comment: if possible, maybe we should rent resources on Amazon to test with larger networks.}
\end{abstract}