### 2017-02-08 report/main.tex
Fix figures and last typos

% !TeX program = lualatex
\documentclass[10pt,conference,a4paper]{IEEEtran}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage[rm={oldstyle=false}]{cfr-lm}
\usepackage[australian,american]{babel}

\usepackage[backend=biber,style=ieee,bibencoding=utf8,sorting=none,doi=false,isbn=false,url=false,date=short]{biblatex}
\usepackage{amssymb}
\usepackage[cmex10]{amsmath}
\usepackage{csquotes}
\usepackage{cleveref}
\usepackage[nolist]{acronym}
\usepackage{float}
\usepackage{todonotes}
\usepackage{pgfplots}
\usepackage{pgf}
\usepackage[binary-units]{siunitx}
\usepackage{pgfplotstable}
\usepackage{booktabs}
\usepackage{multirow}
\usepackage{lipsum}
\usepackage[hidelinks]{hyperref}
\usepackage{url}
\usepackage{xspace}
\usepackage[caption=false]{subfig}

\usetikzlibrary{patterns}

\usepgfplotslibrary{external, groupplots, units}
\tikzexternalize
\tikzsetexternalprefix{generated-figures/}

\sisetup{
	round-mode          = places,
	round-precision     = 2,
}
\DeclareSIUnit\mbps{MB/s}
\DeclareSIUnit\gbps{Gb/s}

\pgfplotscreateplotcyclelist{my colors}{%
solid, color=black\\%
densely dotted, color=blue\\%
densely dashed, color=red\\%
}
\makeatother

%% Environment for comments: Set the boolean to false to produce a comment-free version.
\newboolean{showcomments}
\setboolean{showcomments}{false}
\ifthenelse{\boolean{showcomments}}
{ \newcommand{\mynote}[3]{
		\fbox{\bfseries\sffamily\scriptsize#1}
		{\small$\blacktriangleright$\textsf{\textit{\color{#3}{#2}}}$\blacktriangleleft$}}}
{ \newcommand{\mynote}[3]{}}
% One command per author:
\newcommand{\jt}[1]{\mynote{Jocelyn}{#1}{blue}}
\newcommand{\hm}[1]{\mynote{Hugues}{#1}{orange}}
\newcommand{\mm}[1]{\mynote{Miguel}{#1}{gray}}

\newcommand{\epto}{\textsc{EpTO}\xspace}
\newcommand{\jgroups}{\textsc{JGroups}\xspace}
\newcommand{\sys}{\textsc{LSDSuite}\xspace}

\newif\ifjmcs
% Change false to true to have the JMCS cover page
\jmcstrue

\pgfplotsset{
	% compat=1.14,
	tick label style={font=\small},
	label style={font=\small},
	legend style={font=\small},
	/pgfplots/short line legend/.style={
		legend image code/.code={
			\draw[mark repeat=2,mark phase=2,##1]
			plot coordinates {
				(0cm,0cm)
				(0.2cm,0cm)
				(0.4cm,0cm)
			};%
		}
	},
	/pgfplots/xtick pos=left,
}

\def\ieeetitle{EpTOTester}
\def\ieeeauthor{X,X,X}

\def\thesistitle{\epto}
\def\thesissubtitle{Implementation of a Large-Scale Epidemic Total Order Algorithm
}
\def\thesisauthor{Jocelyn Thode}

\author{\IEEEauthorblockN{\ieeeauthor}
	\IEEEauthorblockA{University of Fribourg, Switzerland\\
		\href{mailto:first.last@unifr.ch}{first.last@unifr.ch}}
}
\title{\ieeetitle}

\addbibresource{references.bib}
\renewcommand*{\bibfont}{\small}

\hypersetup{
	pdftitle=\thesistitle,
	pdfauthor=\thesisauthor
}

% \noautocite{*}

\begin{document}

\ifjmcs
\begin{titlepage}
	\begin{otherlanguage}{australian}
		\begin{center}
			\begin{figure}[t]
				\center{\includegraphics[scale=0.2]{logos/MSc_quer.png}}
				\vspace{0.4in}
			\end{figure}

			{\bfseries\Huge \thesistitle \par
				\Large \vspace{0.1in} \thesissubtitle \par}

			\vspace{0.3in}
			\LARGE{\textbf{Master Thesis} \\}
			\vspace{0.4in}

			{\Large \thesisauthor}

			\vspace{0.3in}
			{\Large Université de Fribourg \par}
			\vfill

			\vspace{0.3in}
			{\Large \today \par}

			\vspace{0.9in}

			% === Logos ==============================================
			\begin{figure}[htp]
				\centering
				\subfloat{\includegraphics[scale=0.60]{logos/UNI_Bern.pdf}}\hfill
				\subfloat{\includegraphics[scale=0.54]{logos/UNI_Neuchatel.pdf}}\hfill
				\subfloat{\includegraphics[scale=0.81]{logos/UNI_Fribourg.pdf}}
			\end{figure}
			% === // Logos ==========thun=================================

		\end{center}

	\pagebreak

	\section*{Abstract}

	\noindent 
	\epto is one of the recently introduced total order algorithms for large-scale distributed systems, and provides total order and scalability at the same time. In this paper, we verify this claim by designing a framework to facilitate the evaluation of large scale distributed systems, implement \epto and evaluate its reliability in real-world conditions by comparing it to a deterministic total order algorithm provided by the \jgroups toolkit. We first compare them in a perfect environment assessing the scalability in terms of events throughput and number of peers. We then compare them using a synthetic churn trace and a real one using the same metrics. 
	Our results show \epto is performing as expected, even though in the system sizes we tested \jgroups is still more efficient in a stable environment.
	By extrapolating our results for larger system sizes we can clearly see that \jgroups starts to degrade with the system size. 
	Still, further testing on a bigger cluster is required to evaluate how \epto performs when \jgroups is no longer able to keep up with scale.

	\bigskip

	Dr Hugues Mercier, Institutes of Computer Science and Mathematics, University of Neuchâtel, Supervisor

	Dr Miguel Matos, HASLab - INESC TEC \&, University of Minho, Portugal,  Co-supervisor

	\end{otherlanguage}
\end{titlepage}

\pagebreak
\fi

%	\maketitle

\input{abstract.tex}
\input{introduction.tex}
\input{definitions.tex}
\input{background.tex}
\input{framework.tex}
\input{evaluation.tex}
\input{future-work.tex}
\input{conclusion.tex}

\printbibliography

\end{document}